Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-14 csf-2017/paper.tex
stack management

\textbf{Stack management discipline}
The precautions the needed to safely interact with an adversary have now been introduced and motivated. The stack discipline comprises the precautions. Summarized, they were:
\begin{description}[font=\normalfont\itshape]
\item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system does not hand out global write-local capabilities.
\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used (if any part of the stack has been passed to an adversary, then that part must be erased).
\item[Before invoking the adversary] push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in register $r_0$, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
\item[Before invoking an adversary callback] make sure it is a global capability.
\item[When called by an adversary] make sure the alleged stack pointer has permission $\rwlx$.
\end{description}
\textbf{Reusable macro instructions} 
We define a number of reusable macros. Some of them capture the stack
management discipline and others make the examples in
Section~\ref{sec:examples} readable. The macros are:
\texttt{fetch $r$ $\var{name}$} fetches the capability related to
$\var{name}$ from the linking table and stores it in register
$r$. \texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
macros that uses the stack assumes a stack pointer in register
$r_\stk$.  \texttt{prepstk $r$} is used when a callback is invoked by
the adversary and prepares the received stack pointer by checking that
it has permission $\rwlx$ and making it conform to our stack
discipline. \texttt{scall
  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} Jumps
to the capability in register $r$ in the manner described above. That
is, it pushes local state (the contents of registers
$\overline{r_{\var{priv}}}$), stack pointer and return pointer to the
stack, sets up the necessary on stack return code and creates a return
pointer, restricts the stack pointer, clears the unused part of the
stack, clears the necessary registers and jumps to $r$. Upon return,
the private state is restored. \texttt{mclear $r$} clears all the
memory the capability in register $r$ has authority
over. \texttt{rclear $\var{regSet}$} clears all the registers in
$\var{regSet}$. \texttt{reqglob $r$} checks whether the word in
register $r$ is a global capability. \texttt{crtcls
  $\overline{(x_i,r_i)}$ $r$} allocates a closure pair where the
capability in register $r$ is used to point to the program and a new
environment is allocated (using malloc) where the contents of
$\overline{r_i}$ is stored. In the program referred to by $r$, an
implicit fetch happens when an instruction refers to $x_i$.

>>>>

\textbf{Stack management discipline}
The precautions the needed to safely interact with an adversary have now been introduced and motivated. The stack discipline comprises the precautions. Summarized, they were:
\begin{description}[font=\normalfont\itshape]
\item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system does not hand out global write-local capabilities.
\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used (any part of the stack that has been passed to an adversary must also be erased).
\item[Before invoking the adversary] push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in a register, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
\item[Before invoking an adversary callback] make sure it is a global capability.
\item[When called by an adversary] make sure the alleged stack pointer has permission $\rwlx$.
\end{description}
\textbf{Reusable macro instructions} 
We define a number of reusable macros. Some of them capture the stack
management discipline and others make the examples in
Section~\ref{sec:examples} readable. The macros are:
\texttt{fetch $r$ $\var{name}$} fetches the capability related to
$\var{name}$ from the linking table and stores it in register
$r$. \texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
macros that uses the stack assumes a stack pointer in register
$r_\stk$.  \texttt{prepstk $r$} is used when a callback is invoked by
the adversary and prepares the received stack pointer by checking that
it has permission $\rwlx$ and making it conform to our stack
discipline. \texttt{scall
  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} Jumps
to the capability in register $r$ in the manner described above. That
is, it pushes local state (the contents of registers
$\overline{r_{\var{priv}}}$), stack pointer and return pointer to the
stack, sets up the necessary on stack return code and creates a return
pointer, restricts the stack pointer, clears the unused part of the
stack, clears the necessary registers and jumps to $r$. Upon return,
the private state is restored. \texttt{mclear $r$} clears all the
memory the capability in register $r$ has authority
over. \texttt{rclear $\var{regSet}$} clears all the registers in
$\var{regSet}$. \texttt{reqglob $r$} checks whether the word in
register $r$ is a global capability. \texttt{crtcls
  $\overline{(x_i,r_i)}$ $r$} allocates a closure pair where the
capability in register $r$ is used to point to the program and a new
environment is allocated (using malloc) where the contents of
$\overline{r_i}$ is stored. In the program referred to by $r$, an
implicit fetch happens when an instruction refers to $x_i$.
